Learning from Data Through Models

### 2023-07-14 paper/paper.Rmd
regressions and hyothesis tests as scripts

---
title: "Learning from Data Through Models"
date: '`r format(Sys.time(), "%B %d, %Y")`'
author: "Alberto Bisin, Guillaume Frechette and Jimena Galindo"
abstract: "TBW"
#classoption: pagebackref # passes pagebackref=true to hyperref before it is loaded, allowing backreferencing
output: 
    pdf_document:
        citation_package: natbib
        fig_width: 7
        fig_height: 6
        fig_caption: true
        number_sections: true
        keep_tex: true
        extra_dependencies:
            footmisc: ["bottom"] # footnote management
            setspace: ["doublespacing"] # spacing of the paper
            caption: ["normal"]
            dsfont: null # indicator function 1
            booktabs: null
            makecell: null
            hyperref: null
        includes:
          in_header: extra_header.tex

@@

```{r dependencies, include=FALSE}
library(data.table)
library(here)
library(huxtable)
library(broom)
renv::snapshot()
library("languageserver")
library(reticulate)

# Format the numbers in the inline code output
inline_hook <- function(x){
  if(is.numeric(x)) {

@@

```{r, out.width='50%', fig.align='center', fig.cap='\\label{examp} Polarization regions for two single variable models'}
knitr::include_graphics(here("diagram_polariazation_2variables.png"))
```

@@

```{r learningAll, out.width='50%', fig.align='center', fig.cap='\\label{learningAll} Share of correct predictions across rounds by performance in rounds 20 to 40'}
knitr::include_graphics(here("computed_objects", "figures", "learning_all.png"))
```

@@

```{r p2groups, out.width='50%', fig.align='center', fig.cap='\\label{p2groups} Share of correct predictions across rounds'}
knitr::include_graphics(here("computed_objects", "figures", "p2_correct_rounds.png"))
```

@@

In part 2, subjects were allowed to disclose up to $m$ variables before making their predictions. and $m$ was assigned at random 
for every subject in every round. We use this random assignment to understand if allowing them to reveal an additional variable 
improves their prediction accuracy. If it does, it must be that they are using the additional information for prediction. 
Figure \ref{simplicity} shows the share of correct predictions by the number of variables that subjects chose to disclose.

```{r simplicity, out.width='50%', fig.align='center', fig.cap='\\label{simplicity} Share of correct predictions by number of disclosed variables'}
knitr::include_graphics(here("computed_objects", "figures", "revealed_variables_pooled.png"))
```
We see that there is a sharp increase when going from 1 to 2 variables, but no significant increase when going from 2 to 3 and
beyond. This suggests that on average subjects are using mental models that consider 2 variables. However, since subjects were 
allowed to reveal fewer variables than what was assigned to them, the estimates in \ref{simplicity} might be biased. To account 
for this, we look at the share of correct predictions by the number of variables that subjects were allowed to disclose. The results 
are presented in Table

```{r }
m1 <- lm(player.correct ~ C(player.number_variables) - 1, data = part2)
m2 <- lm(player.correct ~ C(revealed_variables_count) - 1, data = part2)

huxreg("Disclosed" = m2, 
       "Allowed" = m1,
       coefs = c("Zero variables" = "C(revealed_variables_count)0",
                 "One variable" = "C(player.number_variables)1",
                 "One variable" = "C(revealed_variables_count)1",
                 "Two variables" = "C(player.number_variables)2",
                 "Two variables" = "C(revealed_variables_count)2",
                 "Three variables" = "C(player.number_variables)3",
                 "Three variables" = "C(revealed_variables_count)3",
                 "Four variables" = "C(player.number_variables)4",
                 "Four variables" = "C(revealed_variables_count)4",
                 "Five variables" = "C(player.number_variables)5",
                 "Five variables" = "C(revealed_variables_count)5"), 
        statistics = c("N. obs." = "nobs", "R squared" = "r.squared" ))

@@

## Model Choices

@@

# Appendix

>>>>

---
title: "Learning from Data Through Models"
date: '`r format(Sys.time(), "%B %d, %Y")`'
author: "Alberto Bisin, Guillaume Frechette and Jimena Galindo"
abstract: "TBW"
#classoption: pagebackref # passes pagebackref=true to hyperref before it is loaded, allowing backreferencing
output: 
    pdf_document:
        citation_package: natbib
        fig_width: 7
        fig_height: 6
        fig_caption: true
        number_sections: true
        template: NULL
        keep_tex: true
        extra_dependencies:
            footmisc: ["bottom"] # footnote management
            setspace: ["doublespacing"] # spacing of the paper
            caption: ["normal"]
            dsfont: null # indicator function 1
            booktabs: null
            makecell: null
            hyperref: null
        includes:
          in_header: extra_header.tex

@@

```{r dependencies, include=FALSE}
source("../parameters_and_packages.R")
```

```{r setup, include=FALSE}
# Format the numbers in the inline code output
inline_hook <- function(x){
  if(is.numeric(x)) {

@@

```{r, out.width='65%', fig.align='center', fig.cap='\\label{examp} Polarization regions for two single variable models'}
knitr::include_graphics(here("diagram_polariazation_2variables.png"))
```

@@

```{r learningAll, out.width='65%', fig.align='center', fig.cap='\\label{learningAll} Share of correct predictions across rounds by performance in rounds 20 to 40'}
knitr::include_graphics(here("computed_objects", "figures", "learning_all.png"))
```

@@

```{r p2groups, out.width='65%', fig.align='center', fig.cap='\\label{p2groups} Share of correct predictions across rounds'}
knitr::include_graphics(here("computed_objects", "figures", "p2_correct_rounds.png"))
```

@@

In part 2, subjects were allowed to disclose up to $m$ variables before making their predictions. and $m$ was assigned at random 
for every subject in every round. We use this random assignment to understand if allowing them to reveal an additional variable 
improves their prediction accuracy. If it does, it must be that they are using the additional information for prediction. 
Figure \ref{simplicity} shows the share of correct predictions by the number of variables that subjects chose to disclose. 
Figure \ref{simplicity2} shows the number of variables that subjects chose to disclose by the number they we allowed to disclose.
On average, they always disclose fewer variables than what is available to them. This is a strong indication that at least some
of the subjects are using simple models. 
Further evidence is provided by the accuracy of their predictions as discussed in what follows

```{r simplicity2, out.width='65%', fig.align='center', fig.cap='\\label{simplicity2} Number of disclosed variables by number of allowed variables'}
knitr::include_graphics(here("computed_objects", "figures", "revealed_available.png"))
```

```{r simplicity, out.width='65%', fig.align='center', fig.cap='\\label{simplicity} Share of correct predictions by number of disclosed variables'}
knitr::include_graphics(here("computed_objects", "figures", "revealed_variables_pooled.png"))
```

We see that there is a sharp increase when going from 1 to 2 variables, but no significant increase when going from 2 to 3 and
beyond. This suggests that on average subjects are using mental models that consider 2 variables. However, since subjects were 
allowed to reveal fewer variables than what was assigned to them, the estimates in \ref{simplicity} might be biased. To account 
for this, we also look at the share of correct predictions by the number of variables that subjects were allowed to disclose. The results 
are presented in Table \ref{regression_table}. The table presents the regression estimates for the model with the assigned number of
variables as well as the model with the number of variables that subjects chose to disclose. The last column clusters standard errors
at the group level.

```{r regressionTable}
regs <- readRDS(here("computed_objects", "tables", "regressions.rds"))

regression_table <- huxreg("Disclosed" = regs$m2,
                           "Allowed" = regs$m1,
                           "Allowed" = regs$m4,
       coefs = c("Zero variables" = "C(revealed_variables_count)0",
                 "One variable" = "C(player.number_variables)1",
                 "One variable" = "C(revealed_variables_count)1",
                 "Two variables" = "C(player.number_variables)2",
                 "Two variables" = "C(revealed_variables_count)2",
                 "Three variables" = "C(player.number_variables)3",
                 "Three variables" = "C(revealed_variables_count)3",
                 "Four variables" = "C(player.number_variables)4",
                 "Four variables" = "C(revealed_variables_count)4",
                 "Five variables" = "C(player.number_variables)5",
                 "Five variables" = "C(revealed_variables_count)5"),
        statistics = c("N. obs." = "nobs", "R squared" = "r.squared")
        )

row <- data.frame("Clustered SE", "NO", "NO", "Group")
regression_table <- insert_row(regression_table, row, after = 13)
caption(regression_table) <- "Regression results, dependent variable is the share of correct predictions."
label(regression_table) <- "regression_table"
regression_table
```

```{r hypothesesAll}
# load the tables
tables <- readRDS(here("computed_objects", "tables", "hypothesis_tests.rds"))
results_tbl1 <- tables$tbl1
results_tbl2 <- tables$tbl2
results_tbl4 <- tables$tbl4
# assigned variables
test_assigned <- results_tbl1 %>% as_huxtable() %>% set_col_width(c(0.5, 0.25, 0.25))
label(test_assigned) <- "test_assigned"
caption(test_assigned) <- "Hypothesis tests for the regression with the number of variables that subjects were allowed to disclose."

# clustered standard errors
test_clustered <- results_tbl4 %>% as_huxtable() %>% set_col_width(c(0.5, 0.25, 0.25))
label(test_clustered) <- "test_clustered"
caption(test_clustered) <- "Hypothesis tests for the regression with the number of variables that subjects were allowed to disclose and clustered standard errors."

# revealed variables
test_revealed <- results_tbl2 %>% as_huxtable() %>% set_col_width(c(0.5, 0.25, 0.25))
label(test_revealed) <- "test_revealed"
caption(test_revealed) <- "Hypothesis tests for the regression with the number of variables that subjects chose to disclose."

@@

We are interested in testing whether the probability of guessing the state correctly increases when subjects are allowed to 
disclose an additional variable. This corresponds to testing the linear hypotheses $\beta_{m+1}-\beta_m=0$ for $m=1, 2, 3, 4$.
The results of these tests are presented in Table \ref{hypotheses}. We see that the probability of guessing the state when subjects
are allowed to disclose 2 variables is significantly higher than when they are allowed to disclose only 1 variable. And the effect 
of having access to the additional variable is `r results_tbl4[2, Estimate]` percentage points. Allowing 3, 4 and 5 variables has no 
effect, which sugests that subjects are using models that consider 2 variables. All the hypothesis tests are reported in 
Tables \ref{test_assigned},\ref{test_revealed} and \ref{test_clustered} in the appendix.

## Model Choices
Since we now know that subjects are using models that consider 2 variables, we can look at the specific models that they are using.
We look at two particular features: Are they choosing the most informative models? and do they improve their model choices over time?

@@

\newpage
\appendix

# Appendix
```{r appendix}
test_revealed
test_assigned
test_clustered
```
